---
title: "Nonlinear Regression (Archontoulis and Miguez) paper"
output: html_notebook
---

<https://femiguez.github.io/nlraa-docs/nlraa-Oddi-LFMC.html>

<https://onlinelibrary.wiley.com/doi/pdfdirect/10.1002/ece3.5543>

```{r}
# install.packages("emmeans")

library(nlme)
library(nlraa)
library(ggplot2)
library(emmeans) # emmeans::contrast(), emmeans::emmeans()
library(tidyverse)
```

# Steps in fitting a nonlinear mixed model

## Part I.

```{r}
data(lfmc)
glimpse(lfmc)
```

```{r}
lfmc
```

```{r}
lfmc %>%
  count(leaf.type)

lfmc %>%
  count(time)

lfmc %>%
  count(plot)


lfmc %>%
  count(site)

lfmc %>%
  count(group)

lfmc %>%
  count(site, plot)
```

```{r}
lfmc %>%
  ggplot(aes(time, lfmc)) +
  geom_point() +
  facet_wrap(~group)
```

```{r}
lfmc.gd <- groupedData(lfmc ~ time | group, data = lfmc, order.groups = FALSE)
lfmc.gd
```

```{r}
lfmc.gd$plot <- with(lfmc.gd, factor(plot, levels = c("4", "5", "6", "1", "2", "3")))
lfmc.gd
```

```{r}
ggplot(lfmc, aes(time, lfmc)) +
  geom_point() +
  geom_smooth(method = "loess") +
  facet_wrap(~leaf.type) +
  xlab("Time (days)") +
  ylab("LMFC (%)")
```

## NONLINEAR MIXED-EFFECTS MODEL (M1)

```{r}
nlsL <- nlsList(lfmc ~ SSdlf(time, A, w, m, s), data = lfmc.gd)
coef(nlsL)
```

```{r, fig.width = 4, height = 3}
plot(intervals(nlsL))
```

```{r}
nl.re <- nlme(nlsL, control = nlmeControl(maxIter = 5000, msMaxIter = 1500))
```

```{r}
nl.re.1 <- update(nl.re, random = pdDiag(A + w + m + s ~ 1))
fxf <- fixef(nl.re.1)
fxf
```

```{r}
# nl.me <- update(nl.re.1, fixed = list(A + w + m + s ~ leaf.type),
#                 start = c(A = fxf[1],0,0,0,
#                           w = fxf[2],0,0,0,
#                           m = fxf[3],0,0,0,
#                           s = fxf[4],0,0,0))
```

```{r}
nl.re.2 <- update(nl.re.1, random = pdDiag(A + w + m ~ 1))
nl.re.2
```

```{r}
fxf <- fixef(nl.re.2)
fxf
```

```{r}
nl.me.1 <- update(
  nl.re.2,
  fixed = list(A + w + m + s ~ leaf.type),
  start = c(A = fxf[1], 0, 0, 0, w = fxf[2], 0, 0, 0, m = fxf[3], 0, 0, 0, s = fxf[4], 0, 0, 0)
)
nl.me.1
```

```{r}
AIC(nl.re.2, nl.me.1)
```

```{r}
IC_tab(nl.re.2, nl.me.1)
```

The small StdDev of "w" suggests to remove the random-effect of plot on this parameter.

```{r}
nl.me.2 <- update(nl.me.1, random = pdDiag(A + m ~ 1))
IC_tab(nl.me.1, nl.me.2)
```

in addition, if the random-effects on "m" is also removed.

```{r}
nl.me.3 <- update(nl.me.2, random = pdDiag(A ~ 1))
nl.me.3
```

the model fit is similar, supporting the simpler model.

```{r}
IC_tab(nl.me.2, nl.me.3)
```

```{r}
plot(nl.me.3)
```

```{r}
hist(residuals(nl.me.3, type = "normalized"), breaks = 20)
```

```{r}
rse <- residuals(nl.me.3) / sigma(nl.me.3)
par(pty = "s")
qqnorm(rse)
qqline(rse)
```

Heterocedasticity: variance modeling ----

```{r}
nl.me.3.vm <- update(nl.me.3, weights = varIdent(form = ~ 1 | leaf.type))
nl.me.3.vm
```

```{r}
IC_tab(nl.me.3, nl.me.3.vm)
```

여기서 가중잔차는 weight를 이용해서 계산해야 제대로 plot이 됨

```{r}
res <-
  tibble(
    fitted = fitted(nl.me.3.vm),
    resid = resid(nl.me.3.vm), # 일반 잔차
    nresid = resid(nl.me.3.vm, type = "normalized"), # 표준 잔차
    weight = diag(var_cov(nl.me.3.vm)), # 가중치
    wresid = sqrt(1 / weight) * resid # 가중잔차 수동계산
  ) %>%
  bind_cols(lfmc.gd)
res
```

```{r}
res %>%
  ggplot(aes(fitted, nresid)) +
  geom_point(aes(color = leaf.type))
res %>%
  ggplot(aes(fitted, wresid)) +
  geom_point(aes(color = leaf.type))
plot(nl.me.3.vm)
```

```{r}
hist(residuals(nl.me.3.vm, type = "normalized"), main = "", xlab = "Residuals")
```

the residuals look fine now.

```{r}
qqnorm(resid(nl.me.3.vm, type = "normalized"))
qqline(resid(nl.me.3.vm, type = "normalized"))
```

Temporal correlation check. This plot suggests there is no need for modeling the correlation of the residuals.

```{r}
plot(ACF(nl.me.3.vm), alpha = 0.01)
```

### Evaluation of Fixed Effects

## the model is fitted by Maximum likelihood

```{r}
nl.me.ml <- update(nl.me.3.vm, method = "ML")
nl.me.ml
```

the model is reduced removing the fixed-effect of "leaf type" on "s"

```{r}
nl.me.ml.1 <- update(nl.me.ml,
  fixed = list(A + w + m ~ leaf.type, s ~ 1),
  start = c(A = fxf[1], 0, 0, 0, w = fxf[2], 0, 0, 0, m = fxf[3], 0, 0, 0, s = fxf[4])
)
nl.me.ml.1
```

the AIC comparison suggests that the more complex model is not justified (fits are similar)

```{r}
IC_tab(nl.me.ml, nl.me.ml.1)
```

the model is reduced removing the fixed-effect of "leaf type" on "m"

```{r}
nl.me.ml.2 <- update(nl.me.ml,
  fixed = list(A + w ~ leaf.type, m + s ~ 1),
  start = c(A = fxf[1], 0, 0, 0, w = fxf[2], 0, 0, 0, m = fxf[3], s = fxf[4])
)
nl.me.ml.2
```

the model is reduced removing the fixed-effect of "leaf type" on "W".

```{r}
nl.me.ml.3 <- update(nl.me.ml,
  fixed = list(A ~ leaf.type, w + m + s ~ 1),
  start = c(A = fxf[1], 0, 0, 0, w = fxf[2], m = fxf[3], s = fxf[4])
)
nl.me.ml.3
```

```{r}
nl.me.ml.4 <- update(nl.me.ml,
  fixed = list(w ~ leaf.type, A + m + s ~ 1),
  start = c(A = fxf[1], w = fxf[2], 0, 0, 0, m = fxf[3], s = fxf[4])
)
nl.me.ml.4
```

```{r}
IC_tab(nl.me.ml.1, nl.me.ml.2, nl.me.ml.3, nl.me.ml.4)
```

### Final Model

This model assumes:

-   "A" and "w" vary with leaf type (fixed-effects).
-   "A" varies with plot (random-effects).
-   "m" and "s" are unique for all the plots and leaf types.
-   Residual variance depends on "leaf type"
-   There is no temporal dependence
-   Residuals following a normal distribution

```{r, paged.print = FALSE}
M1 <- update(nl.me.ml.2, method = "REML")
summary(M1)
```

```{r}
fixef(M1)
```

```{r}
ranef(M1)
```

```{r}
intervals(M1)
```

# Part II

```{r}
# 2.2.1 - Overall predictions (Table 3) ----
Agw <- fixef(M1)[1] # "A" for grasses in the W site (GW)
Age <- fixef(M1)[1] + fixef(M1)[2] # "A" for grasses in the E site (GE)
Asm <- fixef(M1)[1] + fixef(M1)[3] # "A" for Mullinum spinosum (SM)
Ass <- fixef(M1)[1] + fixef(M1)[4] # "A" for Senecio filaginoides (SS)
Wgw <- fixef(M1)[5] # "w" for grasses in the W site (GW)
Wge <- fixef(M1)[5] + fixef(M1)[6] # "w" for grasses in the E site (GE)
Wsm <- fixef(M1)[5] + fixef(M1)[7] # "w" for Mullinum spinosum (SM)
Wss <- fixef(M1)[5] + fixef(M1)[8] # "w" for Senecio filaginoides (SS)
M <- fixef(M1)[9] # "m" for all the leaf types
S <- fixef(M1)[10] # "s" for all the leaf types

GW <- subset(lfmc, leaf.type == "Grass W")
GE <- subset(lfmc, leaf.type == "Grass E")
SM <- subset(lfmc, leaf.type == "M. spinosum")
SS <- subset(lfmc, leaf.type == "S. bracteolactus")

par(mfrow = c(2, 2), cex = 0.75)
# Grasses W
plot(GW$time, GW$lfmc, xlab = "", ylab = "LFMC (%)", ylim = c(0, 100), main = "Grasses W site", col = "darkgreen")
curve(((Agw - Wgw) / (1 + exp((M - x) / S)) + Wgw), lwd = 2, lty = 1, add = T, col = "darkgreen")
# Grasses E
plot(GE$time, GE$lfmc, xlab = "", ylab = "LFMC (%)", ylim = c(0, 100), main = "Grasses E site", col = "green")
curve(((Age - Wge) / (1 + exp((M - x) / S)) + Wge), lwd = 2, lty = 1, add = T, col = "green")
# M. Spinosum
plot(SM$time, SM$lfmc, xlab = "Time (days)", ylab = "LFMC (%)", ylim = c(0, 350), main = "M. spinosum", font.main = 4, col = "darkorange")
curve(((Asm - Wsm) / (1 + exp((M - x) / S)) + Wsm), lwd = 2, lty = 1, add = T, col = "darkorange")
# S. filaginoides
plot(SS$time, SS$lfmc, xlab = "Time (days)", ylab = "LFMC (%)", ylim = c(0, 350), main = "S. bracteolactus", font.main = 4, col = "darkred")
curve(((Ass - Wss) / (1 + exp((M - x) / S)) + Wss), lwd = 2, lty = 1, add = T, col = "darkred")
```

```{r}
ranef(M1)
```

```{r}
# 2.2.2 - Predictions at plot level (Table 3) ----
Agw.p4 <- fixef(M1)[1] + ranef(M1)[1, 1] # "A" for grasses in the plot 4 (GW)
Agw.p5 <- fixef(M1)[1] + ranef(M1)[2, 1] # "A" for grasses in the plot 5 (GW)
Agw.p6 <- fixef(M1)[1] + ranef(M1)[3, 1] # "A" for grasses in the plot 6 (GW)
Age.p1 <- fixef(M1)[1] + fixef(M1)[2] + ranef(M1)[4, 1] # "A" for grasses in the plot 1 (GW)
Age.p2 <- fixef(M1)[1] + fixef(M1)[2] + ranef(M1)[7, 1] # "A" for grasses in the plot 2 (GW)
Age.p3 <- fixef(M1)[1] + fixef(M1)[2] + ranef(M1)[10, 1] # "A" for grasses in the plot 3 (GW)
Asm.p1 <- fixef(M1)[1] + fixef(M1)[3] + ranef(M1)[5, 1] # "A" for M. spinosum in the plot 1 (SM)
Asm.p2 <- fixef(M1)[1] + fixef(M1)[3] + ranef(M1)[8, 1] # "A" for M. spinosum in the plot 2 (SM)
Asm.p3 <- fixef(M1)[1] + fixef(M1)[3] + ranef(M1)[11, 1] # "A" for M. spinosum in the plot 3 (SM)
Ass.p1 <- fixef(M1)[1] + fixef(M1)[4] + ranef(M1)[6, 1] # "A" for S. filaginoides in the plot 1 (SM)
Ass.p2 <- fixef(M1)[1] + fixef(M1)[4] + ranef(M1)[9, 1] # "A" for S. filaginoides in the plot 2 (SM)
Ass.p3 <- fixef(M1)[1] + fixef(M1)[4] + ranef(M1)[12, 1] # "A" for S. filaginoides in the plot 3 (SM)

par(mfrow = c(2, 2), cex = 0.75) # (Figure 4)
# Grasses W site
plot(GW$time, GW$lfmc, xlab = "", ylab = "LFMC (%)", ylim = c(0, 100), main = "Grasses W site", col = "darkgreen")
curve(((Agw.p4 - Wgw) / (1 + exp((M - x) / S)) + Wgw), lwd = 1, lty = 2, add = T, col = "darkgreen") # plot 4
curve(((Agw.p5 - Wgw) / (1 + exp((M - x) / S)) + Wgw), lwd = 1, lty = 2, add = T, col = "darkgreen") # plot 5
curve(((Agw.p6 - Wgw) / (1 + exp((M - x) / S)) + Wgw), lwd = 1, lty = 2, add = T, col = "darkgreen") # plot 6
# Grasses E site
plot(GE$time, GE$lfmc, xlab = "", ylab = "LFMC (%)", ylim = c(0, 100), main = "Grasses E site", col = "green")
curve(((Age.p1 - Wge) / (1 + exp((M - x) / S)) + Wge), lwd = 1, lty = 2, add = T, col = "green") # plot 1
curve(((Age.p2 - Wge) / (1 + exp((M - x) / S)) + Wge), lwd = 1, lty = 2, add = T, col = "green") # plot 2
curve(((Age.p3 - Wge) / (1 + exp((M - x) / S)) + Wge), lwd = 1, lty = 2, add = T, col = "green") # plot 3
# M. Spinosum (E site)
plot(SM$time, SM$lfmc, xlab = "Time (days)", ylab = "LFMC (%)", ylim = c(0, 350), main = "M. spinosum", font.main = 4, col = "darkorange")
curve(((Asm.p1 - Wsm) / (1 + exp((M - x) / S)) + Wsm), lwd = 1, lty = 2, add = T, col = "darkorange") # plot 1
curve(((Asm.p2 - Wsm) / (1 + exp((M - x) / S)) + Wsm), lwd = 1, lty = 2, add = T, col = "darkorange") # plot 2
curve(((Asm.p3 - Wsm) / (1 + exp((M - x) / S)) + Wsm), lwd = 1, lty = 2, add = T, col = "darkorange") # plot 3
# S. filaginoides (E site)
plot(SS$time, SS$lfmc, xlab = "Time (days)", ylab = "LFMC (%)", ylim = c(0, 350), main = "S. bracteolactus", font.main = 4, col = "darkred")
curve(((Ass.p1 - Wss) / (1 + exp((M - x) / S)) + Wss), lwd = 1, lty = 2, add = T, col = "darkred") # plot 1
curve(((Ass.p2 - Wss) / (1 + exp((M - x) / S)) + Wss), lwd = 1, lty = 2, add = T, col = "darkred") # plot 2
curve(((Ass.p3 - Wss) / (1 + exp((M - x) / S)) + Wss), lwd = 1, lty = 2, add = T, col = "darkred") # plot 3
```

```{r}
# 2.2.3 - Derivatives (drying speed) ----
lfmc.GW <- expression((Agw - Wgw) / (1 + exp((M - x) / S)) + Wgw) # LFMC dynamics for grasses in the W site
ds.GW <- deriv(lfmc.GW, "x", function.arg = TRUE) # Drying speed for grasses in the W site

lfmc.GE <- expression((Age - Wge) / (1 + exp((M - x) / S)) + Wge) # LFMC dynamics for grasses in the E site
ds.GE <- deriv(lfmc.GE, "x", function.arg = TRUE) # Drying speed for grasses in the E site

lfmc.SM <- expression((Asm - Wsm) / (1 + exp((M - x) / S)) + Wsm) # LFMC dynamics for M. spinosum
ds.SM <- deriv(lfmc.SM, "x", function.arg = TRUE) # Drying speed for M. spinosum

lfmc.SS <- expression((Ass - Wss) / (1 + exp((M - x) / S)) + Wss) # LFMC dynamics for S. filaginoides
ds.SS <- deriv(lfmc.SS, "x", function.arg = TRUE) # # Drying speed for S. filaginoides

xvec <- seq(0, 100, length = 1000)
y.ds.GW <- ds.GW(xvec)
y.ds.GE <- ds.GE(xvec)
y.ds.SM <- ds.SM(xvec)
y.ds.SS <- ds.SS(xvec)
```

```{r}
par(mfrow = c(2, 2), cex = 0.75) # (Figure 4)
plot(xvec, y.ds.GW,
  xlim = c(0, 80), ylim = c(0, 4),
  ylab = "Drying speed", xlab = "Time (days)", type = "n", main = "Grasses in the W site"
)
lines(xvec, -1 * (attr(y.ds.GW, "grad")), lwd = 2, lty = 1, col = "darkgreen")
plot(xvec, y.ds.GE,
  xlim = c(0, 80), ylim = c(0, 4),
  ylab = "Drying speed", xlab = "Time (days)", type = "n", main = "Grasses in the E site"
)
lines(xvec, -1 * (attr(y.ds.GE, "grad")), lwd = 2, lty = 1, col = "green")
plot(xvec, y.ds.SM,
  xlim = c(0, 80), ylim = c(0, 4),
  ylab = "Drying speed", xlab = "Time (days)", type = "n", main = "M. spinosum", font.main = 4
)
lines(xvec, -1 * (attr(y.ds.SM, "grad")), lwd = 2, lty = 1, col = "darkorange")
plot(xvec, y.ds.SS,
  xlim = c(0, 80), ylim = c(0, 4),
  ylab = "Drying speed", xlab = "Time (days)", type = "n", main = "S. filaginoides", font.main = 4
)
lines(xvec, -1 * (attr(y.ds.SS, "grad")), lwd = 2, lty = 1, col = "darkred")
```

```{r}
# 2.3 ---- Testing differences among leaf types -------------------------------------------

# 2.3.1 - Parameter "A" ----
contrast(emmeans(M1, ~leaf.type, param = "A"), "pairwise")
```

-   Maximum LFMC in grasses from the W site is lower than grasses from the E site (Grass W - Grass E)
-   Maximum LFMC in grasses is lower than shrubs (Grass W - M. spinosum),
-   Maximum LFMC in grasses is lower than shrubs (Grass W - S. bracteolactus)
-   Maximum LFMC in grasses is lower than shrubs (Grass E - M. spinosum)
-   Maximum LFMC in grasses is lower than shrubs (Grass E - S. bracteolactus)
-   There are not difference between the two shrub species (M. spinosum - S. bracteolactus).

```{r}
# 2.3.2 - Parameter "w" ----
contrast(emmeans(M1, ~leaf.type, param = "w"), "pairwise")
```

-   Minimum LFMC in grasses from the W site is higher than grasses from the E site (Grass W - Grass E)
-   Minimum LFMC in grasses is lower than shrubs (Grass W - M. spinosum)
-   Minimum LFMC in grasses is lower than shrubs (Grass W - S. bracteolactus)
-   Minimum LFMC in grasses is lower than shrubs (Grass E - M. spinosum)
-   Minimum LFMC in grasses is lower than shrubs (Grass E - S. bracteolactus)
-   There are not difference between the two shrub species (M. spinosum - S. bracteolactus).

## 3) NONLINEAR FIXED-EFFEC

Response function: lfmc = (A - w) / (1 + exp((m - time)/s))) + w

```{r}
# 3.1.1 - Base nonlinear model ----
nl.fe.0 <- nls(lfmc ~ ((A - w) / (1 + exp((m - time) / s))) + w,
  start = c(A = 15, m = 30, s = -17, w = 5),
  data = lfmc
)
# here, time is the only predictor variable
nl.fe.0
```

```{r}
b <- coef(nl.fe.0) # coefficients of the base model
b
```

```{r}
# 3.1.2 - Leaf type fixed effect ----
nl.fe.1 <- nls(lfmc ~ ((A[leaf.type] - w[leaf.type]) / (1 + exp((m - time) / s))) + w[leaf.type],
  start = list(A = rep(b[1], 4), m = 25, s = -10, w = rep(b[4], 4)),
  data = lfmc
) # and fit is made using the base model's coefficients as start values
nl.fe.1
```

```{r}
b1 <- coef(nl.fe.1) # coefficients of the model with leaf type and time as predictor variables
b1
```

```{r}
# 3.1.3 - Plot fixed effect ----
nl.fe.2 <- nls(lfmc ~ ((A[group] - w[group]) / (1 + exp((m - time) / s))) + w[group],
  start = list(A = rep(c(b1[1], b1[2], b1[3], b1[4]), 3), m = 25, s = -10, w = rep(c(b1[7], b1[8], b1[9], b1[10]), 3)),
  data = lfmc
) # the model is fit using "b1" as the start values
nl.fe.2
```

```{r}
IC_tab(nl.fe.0, nl.fe.1, nl.fe.2)
```

```{r}
# Residual checking:
par(mfrow = c(1, 1))
plot(nl.fe.2) # heterocedasticity in the residuals
```

```{r}
hist(resid(nl.fe.2), main = "", xlab = "Residuals") # slightly skew distribution
```

```{r}
qqnorm(resid(nl.fe.2))
qqline(resid(nl.fe.2))
```

This model assumes:

-   "A" and "w" vary with leaf type and plot (fixed-effects).
-   "m" and "s" are unique for all the plots and leaf types.
-   Variance is homogeneous
-   There is no temporal dependence
-   Residuals following a normal distribution

```{r}
M2 <- nl.fe.2
summary(M2)
```

## 4 - LINEAR MIXED-EFFECTS MODEL (M3)

Response function: lfmc = β0 + β1xTime + β2xGE + β3xSM + β4xSS β5xGExTime + β6xSMxTime + β7xSSxTime

where GE, SM, and SS are dummy variables (0 or 1) created to indicate differences between the base level of "leaf type", i.e., grasses from the W site [GW], and the remaining levels.

### 4.1.1 Base linear mixed-effect mode

```{r}
# 4.1.1 - Base linear mixed-effect model ----
l.me <- lme(lfmc ~ time * leaf.type, random = ~ 1 | plot, data = lfmc)

# Residual checking:
plot(l.me) # heterocedasticity in the residuals
```

```{r}
plot(ACF(l.me, resType = "normalized", na.action = na.omit), alpha = 0.05, grid = TRUE) # temporal correlation is observed at lag 1
```

```{r}
# 4.1.2 - Variance modelling ----
l.me.vm <- update(l.me, weights = varIdent(form = ~ 1 | leaf.type))

# Residual checking:
plot(l.me.vm) # variances there seem to be well modeled
```

```{r}
AIC(l.me, l.me.vm) # modeling the variance improves the model fit
```

```{r}
# 4.1.3 - Temporal correlation modelling ----
l.me.vm.tm <- update(l.me.vm, correlation = corARMA(p = 1, q = 1, form = ~1))
plot(ACF(l.me.vm.tm, resType = "normalized", na.action = na.omit), alpha = 0.05, grid = TRUE) # the temporal dependence is modeled
```

```{r}
AIC(l.me.vm, l.me.vm.tm) # modeling the temporal correlation improves the model fit
```

```{r}
plot(l.me.vm.tm)
```

```{r}
plot(ACF(l.me.vm.tm, resType = "normalized", na.action = na.omit), alpha = 0.05, grid = TRUE)
```

```{r}
hist(resid(l.me.vm.tm, type = "normalized"), main = "", xlab = "Residuals")
```

```{r}
qqnorm(resid(l.me.vm.tm, type = "normalized"))
qqline(resid(l.me.vm.tm, type = "normalized"))
```

```{r}
# 4.1.4 - Evaluation of the fixed-effects ----
l.me.ml <- update(l.me.vm.tm, method = "ML") # the model is fitted by Maximum likelihood
l.me.ml.1 <- update(l.me.ml, ~ . - time:leaf.type) # the model is reduced removing the interaction "time x leaf type"
IC_tab(l.me.ml, l.me.ml.1) # the AIC comparison suggests the interaction term to be important
```

```{r, paged.print=FALSE}
M3 <- l.me.vm.tm
summary(M3)
```

```{r}
M4 <- lm(lfmc ~ time * group, data = lfmc)
summary(M4)
```

```{r}
# Residual checking:
par(mfrow = c(2, 2))
plot(M4) # a clear pattern in the residuals is observed (model assumptions are not met)
```

```{r}
# 6 - NULL MODEL (M5) ######################################################################################

M5 <- lm(lfmc ~ 1, data = lfmc)
summary(M5)
```

```{r}
IC_tab(M2, M3, M4, M5) # , weights = T, delta = TRUE, base=T, sort = TRUE) # the nonlinear mixed-effects model is clearly the best
```

```{r}
IC_tab(M1)
```

```{r}
par(mfrow = c(2, 2), cex = 0.65) # (Figure 5)
v1 <- c(1, 2, 3, 4)
v2 <- c("GW", "GE", "SM", "SS")

# Nonlinear mixed-effects model (M1)
rM1 <- resid(M1, type = "normalized")
fM1 <- fitted(M1)
plot(fM1, rM1, xlab = "Fitted LFMC (%)", ylab = "Residuals", ylim = c(-3, 3))
abline(a = 0, b = 0, lwd = 2)
boxplot(rM1 ~ lfmc$leaf.type, ylab = "Residuals", xlab = "Leaf type", xaxt = "n", ylim = c(-3, 3))
axis(side = 1, at = v1, labels = v2, las = 1)
plot(rM1 ~ lfmc$time, ylab = "Residuals", xlab = "Time (days)", ylim = c(-3, 3))
abline(a = 0, b = 0, lw = 2)
qqnorm(rM1, main = "", xlab = "Theoretical quantiles", ylab = "Sample quantiles")
qqline(rM1)
```

```{r}
# Nonlinear fixed-effects model (M2)
par(mfrow = c(2, 2), cex = 0.65) # (Figure 5)

rM2 <- resid(M2)
fM2 <- fitted(M2)
plot(fM2, rM2, xlab = "Fitted LFMC (%)", ylab = "Residuals")
abline(a = 0, b = 0, lwd = 2)
boxplot(rM2 ~ lfmc$leaf.type, ylab = "Residuals", xlab = "Leaf type", xaxt = "n")
axis(side = 1, at = v1, labels = v2, las = 1)
plot(rM2 ~ lfmc$time, ylab = "Residuals", xlab = "Time (days)")
abline(a = 0, b = 0, lw = 2)
qqnorm(rM2, main = "", xlab = "Theoretical quantiles", ylab = "Sample quantiles")
qqline(rM2)
```

```{r}
par(mfrow = c(2, 2), cex = 0.65)

# Linear mixed-effects model (M3)
rM3 <- resid(M3, type = "normalized")
fM3 <- fitted(M3)
plot(fM3, rM3, xlab = "Fitted LFMC (%)", ylab = "Residuals", ylim = c(-3, 3))
abline(a = 0, b = 0, lwd = 2)
boxplot(rM3 ~ lfmc$leaf.type, ylab = "Residuals", xlab = "Leaf type", xaxt = "n", ylim = c(-3, 3))
axis(side = 1, at = v1, labels = v2, las = 1)
plot(rM3 ~ lfmc$time, ylab = "Residuals", xlab = "Time (days)", ylim = c(-3, 3))
abline(a = 0, b = 0, lw = 2)
qqnorm(rM3, main = "", ylab = "Sample quantiles", xlab = "Theoretical quantiles")
qqline(rM3)
```

```{r}
par(mfrow = c(2, 2), cex = 0.65)

# Clasical regression (M4)
rM4 <- resid(M4)
fM4 <- fitted(M4)
plot(fM4, rM4, xlab = "Fitted LFMC (%)", ylab = "Residuals")
abline(a = 0, b = 0, lwd = 2)
boxplot(rM4 ~ lfmc$leaf.type, ylab = "Residuals", xlab = "", xaxt = "n")
axis(side = 1, at = v1, labels = v2, las = 1)
plot(rM4 ~ lfmc$time, ylab = "Residuals", xlab = "Time (days)")
abline(a = 0, b = 0, lw = 2)
qqnorm(rM4, main = "", xlab = "Theoretical quantiles", ylab = "Sample quantiles")
qqline(rM4)
```

# ----
