---
title: "Nonlinear Regression with R"
output: html_notebook
---


# setup

```{r}
# devtools::install_github("cran/drc")
# devtools::install_github("cran/alr3")
# devtools::install_github("cran/NRAIA")
# devtools::install_github("cran/nlrwr")
library(nlrwr)
```


# 1 Introduction

## 1.1 A stock-recruitment model

```{r}
## Fig. 1.1. 
plot(num.fish ~ spawn.biomass, 
     data = M.merluccius, 
     xlab = "Spawning biomass (1000 tonnes)",
     ylab = "Recruitment (million fish)")
```


## 1.2 Competition between plant biotypes

```{r}
## Fig. 1.2.
plot(biomass ~ x, 
     data = RScompetition, 
     log = "", 
     xlab = Density ~ (plants/m^2), 
     ylab = Biomass ~ of ~ sensitive ~ biotype ~ (g/plant), 
     pch = as.numeric(as.factor(RScompetition$z)))
```


## 1.3 Grouped dose-response data

```{r}
## Fig. 1.3.
xyplot(DryMatter ~ Dose | Herbicide,
       data = S.alba, 
       scales = list(x = list(log = TRUE)),
       ylab = "Dry matter (g/pot)",
       xlab = "Dose (g/ha)")
```


# 2 Getting Started

## 2.2 Getting started with nls()

### 2.2.1 Introducing the data example

```{r}
L.minor
```


```{r}
## Fig. 2.1.
plot(rate ~ conc, data = L.minor,
     ylab = "Uptake rate (weight/h)", 
     xlab = Substrate ~ concentration ~ (mmol ~ m^-3))
```


### 2.2.2 Model fitting

```{r}
L.minor.m1 <- nls(rate ~ Vm * conc/(K + conc), 
                  data = L.minor, 
                  start = list(K = 20, Vm = 120), 
                  trace = TRUE)
```


```{r}
deviance(L.minor.m1)
```


```{r}
d2 <- sum((L.minor$rate -predict(L.minor.m1))^2)
all.equal(deviance(L.minor.m1), d2)
```


```{r}
logLik(L.minor.m1)
```


```{r}
coef(L.minor.m1)
```


```{r}
summary(L.minor.m1)
```


### 2.2.3 Prediction

```{r}
fitted(L.minor.m1)
```


```{r}
concVal <- with(L.minor, seq(min(conc), max(conc), length.out = 10))
predict(L.minor.m1, newdata = data.frame(conc = concVal))
```


### 2.2.4 Making plots

```{r}
## Fig. 2.2.
plot(rate ~ conc, 
     data = L.minor,
     ylim = c(10, 130), 
     ylab = "Uptake rate (weight/h)", 
     xlab = Substrate ~ concentration ~ (mmol ~ m^-3)) 
lines(L.minor$conc, fitted(L.minor.m1))
```


```{r}
## Fig. 2.3.
plot(rate ~ conc, 
     data = L.minor,
     ylim = c(10, 130), 
     ylab = "Uptake rate (weight/h)", 
     xlab = Substrate ~ concentration ~ (mmol ~ m^-3))

concVal <- with(L.minor, seq(min(conc), max(conc), length.out = 100))
lines(concVal, predict(L.minor.m1, newdata = data.frame(conc = concVal)))
abline(h = coef(L.minor.m1)[2], lty = 2)
```


### 2.2.5 Illustrating the estimation

```{r}
L.minor.m1con <- nlsContourRSS(L.minor.m1)
```


```{r}
## Fig. 2.4. 
par(pty = "s")
plot(L.minor.m1con, col = FALSE, nlev = 10)
```


## 2.3 Generalised linear models

```{r}
## this code should be runned in the clean R session
## library(nlrwr) cause error 
# data(L.minor, package = "nlrwr")
# L.minor.m4 <- glm(rate ~ I(1/conc), 
#                   data = L.minor, 
#                   family = gaussian("inverse"))
# summary(L.minor.m4)
```


## Excercises

```{r}
## 2.1
y0_start = mean(RGRcurve[RGRcurve$Day == 0, "RGR"])
y1 = mean(RGRcurve[RGRcurve$Day == 1, "RGR"])
b_start = 1 / log(y1/y0_start)


e2.1 <- nls(RGR ~ y0 * exp(Day/b),
            data = RGRcurve,
            start = list(y0 = y0_start, b = b_start))

plot(RGRcurve$Day, RGRcurve$RGR, xlab = "Day", ylab = "RGR")
nd <- with(RGRcurve, seq(min(Day), max(Day), length.out = 101))
lines(nd, predict(e2.1, newdata = data.frame(Day = nd)), col = 2)
```


```{r}
## 2.3
plotfit(L.minor.m1)
```


# 3 Starting Values and Self-starters

## 3.1 Finding starting values

### 3.1.1 Graphical exploration

```{r}
## Fig. 3.1.
# install.packages("NISTnls")
data(Chwirut2, package = "NISTnls")
plot(y ~ x, 
     data = Chwirut2, 
     xlab = "Metal distance", 
     ylab = "Ultrasonic response")
```


```{r}
## Fig. 3.2.
expFct <- function(x, beta1, beta2, beta3) {
  exp(-beta1 * x)/(beta2 + beta3 * x)
}

plot(y ~ x, 
     data = Chwirut2, 
     xlab = "Metal distance", 
     ylab = "Ultrasonic response",
     ylim = c(0, 100)
     )
curve(expFct(x, beta1 = 1, beta2 = 0.01, beta3 = 1), add = TRUE, col = 2)
```


```{r}
## Fig. 3.3.
plot(y ~ x, data = Chwirut2, xlab = "Metal distance",
     ylab = "Ultrasonic response", ylim = c(0, 100))
curve(expFct(x, beta1 = 0.1, beta2 = 0.01, beta3 = 1), add = TRUE, lty = 2,
      col = 2)
curve(expFct(x, beta1 = 0.1, beta2 = 0.01, beta3 = 0.1), add = TRUE, lty = 3,
      col = 3)
curve(expFct(x, beta1 = 0.1, beta2 = 0.01, beta3 = 0.01), add = TRUE, lty = 4,
      col = 4)
curve(expFct(x, beta1 = 0.2, beta2 = 0.01, beta3 = 0.01), add = TRUE, lty = 1,
      col = 6)
```


```{r}
Chwirut2.m1 <- nls(
  y ~ expFct(x, beta1, beta2, beta3),
  data = Chwirut2,
  start = list(beta1 = 0.2, beta2 = 0.01, beta3 = 0.01)
)
summary(Chwirut2.m1)
```


### 3.1.2 Searching a grid

```{r}
## bad starting example
# Chwirut2.m0 <- nls(
#   y ~ expFct(x, beta1, beta2, beta3),
#   data = Chwirut2,
#   start = list(beta1 = 1, beta2 = 0.01, beta3 = 1))
# summary(Chwirut2.m0)

## nlsLM may solve the bad starting
Chwirut2.m0 <- minpack.lm::nlsLM(
  y ~ expFct(x, beta1, beta2, beta3),
  data = Chwirut2,
  start = list(beta1 = 1, beta2 = 0.01, beta3 = 1))
summary(Chwirut2.m0)
```


```{r}
grid.Chwirut2 <- expand.grid(beta1 = seq(0.1, 1, by = 0.1),
                             beta2 =  0.01,
                             beta3 = seq(0.1, 1, by = 0.1))
grid.Chwirut2
```


```{r}
Chwirut2.m2a <- nls2(
  y ~ expFct(x, beta1, beta2, beta3),
  data = Chwirut2,
  start = grid.Chwirut2,
  algorithm = "brute-force"
)
Chwirut2.m2a
```


```{r}
nls(
  y ~ expFct(x, beta1, beta2, beta3),
  data = Chwirut2,
  start = as.list(coef(Chwirut2.m2a)))  # starting value from Chwirut2.m2a
```


## 3.2 Using self-starter functions

### 3.2.1 Built-in self-starter functions for nls()

```{r}
L.minor.m2 <- nls(rate ~ SSmicmen(conc, Vm, K), data = L.minor)
summary(L.minor.m2)
```


### 3.2.2 Defining a self-starter function for nls()

```{r}
expModel <- function(predictor, b, y0) {
  y0 * exp(predictor/b)
}

expModelInit <- function(mCall, LHS, data) {
  xy <- sortedXyData(mCall[["predictor"]], LHS, data)
  lmFit <- lm(log(xy[, "y"]) ~ xy[, "x"])
  coefs <- coef(lmFit)
  y0 <- exp(coefs[1])
  b <- 1/coefs[2]
  value <- c(b, y0)
  names(value) <- mCall[c("b", "y0")]
  value
}

SSexp2 <- selfStart(expModel, expModelInit, c("b", "y0"))
```


```{r}
with(RGRcurve, SSexp2(Day, 4, 0.2))
```

**Warning:** selfStart caused an ERROR

```{r}
## ERROR
# getInitial(RGR ~ SSexp2(Day, b, y0), data = RGRcurve)
```


```{r}
#RGRcurve.m1 <- nls(RGR ~ SSexp2(Day, b, y0), data = RGRcurve)
#coef(RGRcurve.m1)
```

More examples on defining self-starter functions for nls() are found in Watkins and Venables (2006) or Venables and Ripley (2002a, pp. 216–217).

# 4 More on nls()

## 4.2 Supplying gradient information

### 4.2.1 Manual supply

```{r}
MMfct1 <- function(conc, K, Vm) {
  numer <- Vm * conc
  denom <- K + conc
  mean <- numer/denom
  partialK <- -numer/(denom^2)
  partialVm <- mean/Vm
  attr(mean, "gradient") <- cbind(partialK, partialVm)
  return(mean)
}
MMfct1(1, 2, 3)
```


```{r}
L.minor.mgr1 <- nls(rate ~ MMfct1(conc, K, Vm), 
                    data = L.minor, 
                    start = list(K = 1, Vm = 1),
                    trace = TRUE)
summary(L.minor.mgr1)
```


```{r}
## without derivative
nls(rate ~ Vm * conc/(K + conc), 
    data = L.minor, start = list(K = 1, Vm = 1), 
    trace = TRUE)
```


### 4.2.2 Automatic supply

```{r}
MMfct2 <- deriv(~ Vm * conc / (K + conc), 
                c("K", "Vm"), 
                function(conc, K, Vm) {})
MMfct2
MMfct2(1, 2, 3)
```


```{r}
L.minor.mgr2 <- nls(
  rate ~ MMfct2(conc, + K, Vm), 
  data = L.minor, 
  start = list(K = 20, Vm = 120))
summary(L.minor.mgr2)
```

Ref: Venables와 Ripley(2002a, pp. 214–216)

## 4.3 Conditionally linear parameters

### 4.3.1 nls() using the "plinear" algorithm

```{r}
L.minor.m3 <- nls(rate ~ conc/(K + conc), 
                  data = L.minor, 
                  algorithm = "plinear",
                  start = list(K = 20))
summary(L.minor.m3)
```


### 4.3.2 A pedestrian approach

```{r}
## Fig. 4.1.
data(segreg)
plot(C ~ Temp, data = segreg, 
     xlab = Mean ~ temperature ~ (degree ~ F),
     ylab = "Energy consumption (kWh)")
```


```{r}
profRSS1 <- function(gamma) {
  deviance(lm(C ~ pmax(0, Temp - gamma), data = segreg))
}
profRSS2 <- Vectorize(profRSS1, "gamma")
```


```{r}
l <- lm( C ~ pmax(0, Temp - 40), data = segreg)
summary(l)
```


```{r}
plot(segreg$Temp, segreg$C)
lines(segreg$Temp, predict(l), col = 2)
```


```{r}
## Fig. 4.2. 
plot(profRSS2(Temp) ~ Temp, 
     data = segreg,
     type = "l", 
     xlab = expression(gamma),
     ylab = "Profile RSS")
```


```{r}
## 좀더 부드럽게 만듬
plot(profRSS2(Temp) ~ Temp, 
     data = data.frame(Temp = seq(min(segreg$Temp), 
                                  max(segreg$Temp), 
                                  length.out = 1001)),
     type = "l", 
     xlab = expression(gamma),
     ylab = "Profile RSS")
```


## 4.4 Fitting models with several predictor variables

### 4.4.1 Two-dimensional predictor

```{r}
RScompetition.m1 <- nls(
  biomass ~ a/(1 + b * (x + c * z)),
  data = RScompetition,
  start = list(a = 20, b = 1, c = 1)
)
summary(RScompetition.m1)
```


```{r}
virDensity <- with(RScompetition, x + coef(RScompetition.m1)[3] * z)
virDenVal <- seq(0, max(virDensity), length.out = 100)
biomassVal <- predict(RScompetition.m1, data.frame(x = virDenVal, z = 0))
```


```{r}
## Fig. 4.3. 
plot(biomassVal ~ virDenVal, type = "l", 
     ylab = "Biomass of sensitive biotype (g/plant)",
     xlab = Virtual ~ density ~ (plants/m^2))
with(RScompetition, points(biomass ~ virDensity))
```


### 4.4.2 General least-squares minimisation

```{r}
## Fig. 4.4. 
par(pty = "s")
plot(Q ~ I, data = IQsig)
theta <- 0:360 * (pi/180)
lines(cos(theta), sin(theta))
```


```{r}
IQsig.m1 <- nls(
  ~ ((I - I0)^2 - 2 * gamma * sin(phi) * (I - I0) * (Q - Q0) + 
       gamma * gamma * (Q - Q0)^2) - (rho * gamma * cos(phi))^2, 
  data = IQsig, 
  start = list(I0 = -0.005, gamma = 1, phi = -0.005, Q0 = -0.005, rho = 1))
summary(IQsig.m1)
```


# 5 Model Diagnostics

## 5.2 Checking the mean structure

### 5.2.1 Plot of the fitted regression curve

```{r}
## Fig. 5.1.
plot(p ~ T, data = vapCO, log = "y",
     xlab = "Temperature (K)",
     ylab = "Pressure (Pa)")
```


```{r}
vapCO.m1 <- nls(
  log(p) ~ A - B/(C + T), 
  data = vapCO,
  start = list(A = 10, B = 100, C = -10))
summary(vapCO.m1)
```


```{r}
## Fig. 5.2.
plot(p ~ T, data = vapCO, log = "y",
     xlab = "Temperature (K)",
     ylab = "Pressure (Pa)")
lines(vapCO$T, exp(fitted(vapCO.m1)), col = 2)
```


```{r}
## Fig. 5.3. 
plot(weight ~ conc, data = lettuce,
     xlab = "Concentration (mg/l)",
     ylab = "Biomass (g)", log = "x")
```


### 5.2.2 Residual plots

```{r}
## Fig. 5.4.
plot(fitted(vapCO.m1), residuals(vapCO.m1),
     xlab = "Fitted Values", ylab = "Residuals")
abline(a = 0, b = 0)
```


```{r}
## Fig. 5.5. 
plot(rootl ~ conc, data = ryegrass,
     xlab = "Concentration (mM)",
     ylab = "Root length (cm)")
```


```{r}
ryegrass.m1 <- lm(rootl ~ as.factor(conc), data = ryegrass)
summary(ryegrass.m1)
```


```{r}
ryegrass.m2 <- nls(
  rootl ~ c + (d - c)/(1 + exp(b * +(log(conc) - log(e)))),
  data = ryegrass,
  start = list(b = 1, c = 0.6, d = 8, e = 3)
)
summary(ryegrass.m2)
```


```{r}
anova(ryegrass.m2, ryegrass.m1)
```


```{r}
Q <- -2 * (logLik(ryegrass.m2) - logLik(ryegrass.m1))
df.Q <- df.residual(ryegrass.m2) - df.residual(ryegrass.m1)
1 - pchisq(Q, df.Q)
```


## 5.3 Variance homogeneity

### 5.3.2 Levene’s test

```{r}
with(ryegrass, leveneTest(rootl, as.factor(conc)))
```


```{r}
## Fig. 5.6. 
plot(fitted(vapCO.m1), abs(residuals(vapCO.m1)),
     xlab = "Fitted values", ylab = "Absolute residuals")
```


## 5.4 Normal distribution

```{r}
## Fig. 5.7. 
plot(vapCO.m1)
```


### 5.4.1 QQ plot

```{r}
## Fig. 5.8. 
standardRes <- residuals(ryegrass.m2)/summary(ryegrass.m2)$sigma

par(pty = "s")
qqnorm(standardRes, main = "")
abline(a = 0, b = 1)
```


### 5.4.2 Shapiro-Wilk test

```{r}
shapiro.test(standardRes)
```


## 5.5 Independence

```{r}
## Fig. 5.9. 
plot(residuals(vapCO.m1), c(residuals(vapCO.m1)[-1], NA),
     xlab = "Residuals", ylab = "Lagged residuals")
```


# 6 Remedies for Model Violations

## 6.1 Variance modelling

### 6.1.1 Power-of-the-mean variance model

```{r}
## Fig. 6.1. 
plot(RGR ~ Day, data = RGRcurve,
     xlab = "Time (days)", ylab = "Relative growth rate (%)")
```


```{r}
y0_start = mean(RGRcurve[RGRcurve$Day == 0, "RGR"])
y1 = mean(RGRcurve[RGRcurve$Day == 1, "RGR"])
b_start = 1 / log(y1/y0_start)

RGRcurve.m2 <- gnls(
  RGR ~  y0 * exp(Day/b), 
  data = RGRcurve,
  start = list(y0 = y0_start, b = b_start),
  weights = varPower()
)
summary(RGRcurve.m2)
```


```{r}
RGRcurve.m2g <- gnls(
  RGR ~  y0 * exp(Day/b), 
  data = RGRcurve,
  start = list(y0 = y0_start, b = b_start)
  # weights = varPower() # without weights
)
summary(RGRcurve.m2g )
```


```{r}
AIC(RGRcurve.m2, RGRcurve.m2g )
```


## 6.2 Transformations

### 6.2.1 Transform-both-sides approach

```{r}
## Fig. 6.2. 
plot(recruits ~ spawners, 
     data = sockeye[-12,],
     xlab = "Number of spawners (thousands)",
     ylab = "Number of recruits (thousands)")
```


일단 nls를 이용하여 fitting을 한 후

```{r}
sockeye.m1 <- nls(
  recruits ~ beta1 * spawners * exp(-beta2 * spawners),
  data = sockeye[-12, ],
  start = list(beta1 = 2, beta2 = 0.001)
)
summary(sockeye.m1)
```


```{r}
## Fig. 6.3. 
plot(fitted(sockeye.m1), abs(residuals(sockeye.m1)),
     xlab = "Fitted values", ylab = "Absolute residuals")
```


### 6.2.2 Finding an appropriate transformation

boxcox.nls를 이용하여 양쪽을 변화시키며 피팅을 한다.

```{r}
## Fig. 6.4.
sockeye.m2 <- boxcox.nls(sockeye.m1)
```


```{r}
bcSummary(sockeye.m2)
```


```{r}
coef(summary(sockeye.m1))
coef(summary(sockeye.m2))
```


## 6.3 Sandwich estimators

이 내용은 정확히 이해하지는 못했음.

```{r}
vcov(sockeye.m1)
```


```{r}
sandwich(sockeye.m1)
```


```{r}
coeftest(sockeye.m1)
```


```{r}
coeftest(sockeye.m1, vcov = sandwich)
```


## 6.4 Weighting

### 6.4.1 Decline in nitrogen content in soil

```{r}
exp1
```


```{r}
## Fig. 6.5. 
plot(Nremaining ~ time, data = exp1,
     xlab = "Time (years)", ylab = "Nitroge content (%)")
```


```{r}
exp1.m1 <- nls(Nremaining ~ SSbiexp(time, a1, a2, b1, b2), data = exp1)
exp1.m2 <- nls(Nremaining ~ SSbiexp(time, a1, a2, b1, b2), data = exp1,
               weights = norep/(stdev * stdev))
```


```{r}
weights(exp1.m2)
```


```{r}
coef(summary(exp1.m1))
coef(summary(exp1.m2))
```


```{r}
exp2
```


```{r}
exp2.m1 <- nls(Nremaining ~ SSbiexp(time, a1, a2, b1, b2), data = exp2)
exp2.m2 <- nls(Nremaining ~ SSbiexp(time, a1, a2, b1, b2), data = exp2,
               weights = norep/(stdev^2))
```


```{r}
## Fig. 6.6. 
plot(Nremaining ~ time, data = exp2,
     xlab = "Time (years)", ylab = "Nitrogen content (%)")
timeVal <- with(exp2, seq(min(time), max(time), length.out = 101))
lines(timeVal, predict(exp2.m1, newdata = data.frame(time = timeVal)), 
      lty = 2, col = 2)
lines(timeVal, predict(exp2.m2, newdata = data.frame(time = timeVal)), 
      lty = 3, col = 3)
```


```{r}
coef(summary(exp2.m1))
coef(summary(exp2.m2))
```


```{r}
summary(exp1.m1)
summary(exp1.m2)
```


# 7 Uncertainty, Hypothesis Testing, and Model Selection

## 7.1 Profile likelihood

```{r}
L.minor.m1 <- update(L.minor.m1, trace = FALSE)
L.minor.m1pro <- profile(L.minor.m1)
```


```{r}
## Fig. 7.1. 
plot(L.minor.m1pro)
```


```{r}
## Fig. 7.2. 
plot(L.minor.m1pro, absVal = FALSE)
```


```{r}
confint(L.minor.m1)
```


## 7.2 Bootstrap

```{r}
set.seed(222)
L.minor.m1boot <- nlsBoot(L.minor.m1)
summary(L.minor.m1boot)
```


```{r}
## Fig. 7.3. 
par(mfrow = c(1, 2))
qqnorm(L.minor.m1boot$coefboot[, 1], main = "K")
qqnorm(L.minor.m1boot$coefboot[, 2], main = "Vm")
```


## 7.3 Wald confidence intervals

```{r}
confint2(L.minor.m1)
```


```{r}
confint2(L.minor.m1, level = 0.99)
```


## 7.4 Estimating derived parameters

```{r}
deltaMethod(L.minor.m1, "Vm/(4*K)")
```


## 7.5 Nested models

```{r}
secalonic
```


```{r}
## Fig. 7.4.
plot(rootl ~ dose, data = secalonic,
     xlab = "Dose (mM)", ylab = "Root length (cm)")
```


### 7.5.1 Using t-tests

```{r}
secalonic.m1 <- nls(rootl ~ SSfpl(dose, a, b, c, d), data = secalonic)
summary(secalonic.m1)
```


### 7.5.2 Using F-tests

```{r}
secalonic.m2 <- nls(rootl ~ SSlogis(dose, a, c, d), data = secalonic)
summary(secalonic.m2)
```


```{r}
anova(secalonic.m2, secalonic.m1)
```


## 7.6 Non-nested models

```{r}
M.merluccius
plot(num.fish ~ spawn.biomass, data = M.merluccius)
```

4가지 조금씩 다른 모델을 비교

```{r}
M.merluccius.bh <- nls(
  num.fish ~ spawn.biomass * alpha/(1 + spawn.biomass/k),
  data = M.merluccius, start = list(alpha = 5, k = 50)
)

M.merluccius.de <- nls(
  num.fish ~ spawn.biomass * alpha * (1 - c * spawn.biomass/k)^(1/c),
  data = M.merluccius, start = list(alpha = 4.4, k = 106, c = 0.86)
)

M.merluccius.ri <- nls(
  num.fish ~ spawn.biomass * alpha * exp(-spawn.biomass/k),
  data = M.merluccius, start = list(alpha = 5, k = 50)
)

M.merluccius.sh <- nls(
  num.fish ~ spawn.biomass * alpha/(1 + (spawn.biomass/k)^c),
  data = M.merluccius, start = list(alpha = 3.87, k = 61.72, c = 2.25),
  control = nls.control(maxiter = 100)
)
```

Residual standard error를 계산하면,

```{r}
summary(M.merluccius.bh)$sigma
summary(M.merluccius.de)$sigma
summary(M.merluccius.ri)$sigma
summary(M.merluccius.sh)$sigma
```

AIC를 비교해보면,

```{r}
AIC(M.merluccius.bh)
AIC(M.merluccius.de)
AIC(M.merluccius.ri)
AIC(M.merluccius.sh)
```

Burnham and Anderson (2002, pp. 70–72)은 한 모델이 다른 모델에 비하여 절대적으로 우위에 있기 위해서는 AIC의 차이가 10 이상의 차이가 필요하다는 경험적 법칙을 제시하였다. 따라서 4 모델 사이에서 큰 차이는 없다. 


# 8 Grouped Data

## 8.1 Fitting grouped data models

```{r}
Puromycin
```


```{r}
## Fig. 8.1.
xyplot(rate ~ conc | state, data = Puromycin,
       xlab = "Substrate concentration (ppm)",
       ylab = "Reaction rates\n(counts/min/min)")
```


### 8.1.1 Using nls()

```{r}
Puromycin.m1 <- nls(
  rate ~ Vm[state] * conc/(K[state] + conc),
  data = Puromycin,
  start = list(K = c(0.1, 0.1), Vm = c(200, 200))
)
summary(Puromycin.m1)
```


### 8.1.2 Using gnls()

```{r}
Puromycin.m2 <- gnls(
  rate ~ Vm * conc/(K + conc),
  data = Puromycin,
  start = list(Vm = c(200, 200), K = c(0.1, 0.1)),
  params = list(Vm ~ state - 1, K ~ state - 1)
)
summary(Puromycin.m2)
```

### 8.1.3 Using nlsList()

```{r}
Puromycin.m3 <- nlsList(
  rate ~ SSmicmen(conc, Vm, K) | state,
  data = Puromycin
)
summary(Puromycin.m3)
```


```{r}
Puromycin2 <- groupedData(rate ~ conc | state, data = Puromycin)
Puromycin.m4 <- nlsList(
  rate ~ SSmicmen(conc, a, b), data = Puromycin2
)
summary(Puromycin.m4)
```


## 8.2 Model reduction and parameter models

### 8.2.1 Comparison of entire groups

```{r}
Puromycin.m5 <- nls(
  rate ~ Vm * conc/(K + conc), data = Puromycin,
  start = list(K = 0.1, Vm = 200))
summary(Puromycin.m5)
```


```{r}
anova(Puromycin.m5, Puromycin.m1)
```


```{r}
AIC(Puromycin.m5, Puromycin.m1)
```

### 8.2.2 Comparison of specific parameters

```{r}
Puromycin.m6 <- nls(
  rate ~ Vm * conc/(K[state] + conc), 
  data = Puromycin, 
  start = list(K = c(0.1, 0.1), Vm = 200))
summary(Puromycin.m6)
```


```{r}
anova(Puromycin.m6, Puromycin.m1)
```


```{r}
Puromycin.m7 <- nls(
  rate ~ Vm[state] * conc/(K + conc), 
  data = Puromycin, start = list(K = 0.1, Vm = c(200, 200))
)
summary(Puromycin.m7)
```


```{r}
anova(Puromycin.m7, Puromycin.m1)
```


AIC로 비교해 보아도 Puromycin.m7이 더욱 좋다. 

```{r}
AIC(Puromycin.m1, Puromycin.m5, Puromycin.m6, Puromycin.m7)
```

## 8.3 Common control

```{r}
G.aparine
```


```{r}
## Fig. 8.2. 
xyplot(drymatter ~ dose | as.factor(treatment),
       data = G.aparine,
       xlab = "Dose (g/ha)", ylab = "Dry matter (mg/pot)")
```

이미 데이터에서 treatment 0가 없어졌음. 아마도 데이터를 업데이트 한 것으로 추측됨. 

```{r}
G.aparine$treatment2 <- factor(G.aparine$treatment)
#levels(G.aparine$treatment2) <- c("0", "1", "2")
G.aparine
```


```{r}
G.aparine.m1 <- nls(
  drymatter ~ c[treatment2] + (d - c[treatment2])/(1 + exp(b[treatment2] * (log(dose) - log(e[treatment2])))),
  data = G.aparine, 
  start = list(b = c(2, 2), c = c(500, 100), d = 1000, e = c(50, 100)))
summary(G.aparine.m1)
```


## 8.4 Prediction

```{r}
concValues <- with(Puromycin, seq(min(conc), max(conc, length.out = 11)))
concValues
```


```{r}
stateVal1 <- levels(Puromycin$state)
stateVal1
```


```{r}
csValues1 <- expand.grid(conc = concValues, state = stateVal1)
csValues1
```


```{r}
predict(Puromycin.m7, newdata = csValues1)
```


```{r}
stateVal2 <- factor("untreated", levels = c("treated", "untreated"))
stateVal2
```


```{r}
csValues2 <- data.frame(conc = concValues, state = stateVal2)
csValues2
```


```{r}
predict(Puromycin.m1, newdata = csValues2)
```


## 8.5 Nonlinear mixed models

```{r}
vinclozolin
```


```{r}
## Fig. 8.3.
xyplot(effect ~ conc | exper, data = vinclozolin,
       xlab = expression(paste("Concentration (", mu, "M)")),
       ylab = "Luminescence (LU")
```


```{r}
LL3.formula <- effect ~ d/(1 + exp(b * (log(conc) - log(e))))

vinclozolin.e1.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 10509,  
  start = list(b = 1, d = 1000, e = 0.26))
vinclozolin.e2.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 10821,  
  start = list(b = 1, d = 1000, e = 0.26))
vinclozolin.e3.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 10828,  
  start = list(b = 1, d = 1000, e = 0.26))
vinclozolin.e4.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 10904,  
  start = list(b = 1, d = 2700, e = 0.03))
vinclozolin.e5.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 11023,  
  start = list(b = 1, d = 1000, e = 0.26))
vinclozolin.e6.m <- nls(
  LL3.formula, data = vinclozolin, subset = exper == 11106,  
  start = list(b = 0.5, d = 2600, e = 0.02))
```


```{r, fig.width=3.3, fig.height=3, warning=FALSE}
## Fig. 8.4. 
#par(pty = "s")
plot(effect ~ conc, data = vinclozolin,
     pch = as.numeric(exper), 
     log = "x",
     xlim = c(1e-04, 10), 
     xlab = expression(paste("Concentration(", mu, "M)")), 
     ylab = "Luminescence(LU)")
concVec <- exp(seq(log(1e-04), log(10), length.out = 50))
lines(concVec, predict(vinclozolin.e1.m, data.frame(conc = concVec)), lty = 2)
lines(concVec, predict(vinclozolin.e2.m, data.frame(conc = concVec)), lty = 3)
lines(concVec, predict(vinclozolin.e3.m, data.frame(conc = concVec)), lty = 4)
lines(concVec, predict(vinclozolin.e4.m, data.frame(conc = concVec)), lty = 5)
lines(concVec, predict(vinclozolin.e5.m, data.frame(conc = concVec)), lty = 6)
lines(concVec, predict(vinclozolin.e6.m, data.frame(conc = concVec)), lty = "3313")
```


```{r}
vinclozolin.m1 <- nlme(
  effect ~ d/(1 + exp(b * (log(conc) - log(e)))), 
  data = vinclozolin,
  fixed = list(b ~ 1, d ~ 1, e ~ 1), 
  random = d ~ 1 | exper, 
  start = c(1, 1000, 0.1))
summary(vinclozolin.m1)
```


```{r, fig.width=3.3, fig.height=3, warning=FALSE}
## Fig. 8.5.
#par(pty = "s")
plot(effect ~ conc, data = vinclozolin,
     pch = as.numeric(exper), 
     log = "x",
     xlim = c(1e-04, 10), 
     xlab = expression(paste("Concentration(", mu, "M)")), 
     ylab = "Luminescence(LU)")
concVec <- exp(seq(log(1e-04), log(10), length.out = 50))
lines(concVec, predict(vinclozolin.e1.m, data.frame(conc = concVec)), lty = 2)
lines(concVec, predict(vinclozolin.e2.m, data.frame(conc = concVec)), lty = 3)
lines(concVec, predict(vinclozolin.e3.m, data.frame(conc = concVec)), lty = 4)
lines(concVec, predict(vinclozolin.e4.m, data.frame(conc = concVec)), lty = 5)
lines(concVec, predict(vinclozolin.e5.m, data.frame(conc = concVec)), lty = 6)
lines(concVec, predict(vinclozolin.e6.m, data.frame(conc = concVec)), lty = "3313")
lines(concVec, 
      predict(vinclozolin.m1, newdata = data.frame(conc = concVec), level = 0), 
      lty = 1, lwd = 3)
```


# ----